A boundary integral equation approach to computing eigenvalues of the Stokes operator

### 2019-04-05 doc/draft-01-stokes/04numerical.tex
add barbell plot
update numerical section

@@

Let the boundary be divided into $N_p$
panels.
%
We parameterize panel $j$ as
$\bx_j(t)$, with $t$ ranging over the
interval $[-1,1]$.
%
Each component of $\bx_j$ is taken to be
a polynomial interpolant over the
standard 16th order Legendre nodes on
$[-1,1]$, denoted by $t_n$, so that
the total number of discretization
points is $N=16N_p$.
%
See \cref{fig:panels} for an example
discretization.
%
Another important quantity below is the
arc-length density of a panel, which
we denote by $s_j(t) := |\bx'_j(t)|$.
%
Finally, we denote the set of panels
which are adjacent to panel $i$
by $A(i)$. On a closed curve,
$A(i)$ contains two integers.

The integral kernels of the single and
double layer potentials have weak
singularities of the form $r^p\log r$
which require special quadrature rules to
achieve high-order accuracy.
%
In the examples below, we use generalized
Gaussian quadrature (GGQ)~\cite{bremer2010}.
%
To demonstrate the idea, we consider
evaluating the convolution of a kernel
$K$ with a density $\sigma$
at the boundary node $\xx_j(t_l)$.
%
GGQ is a Nystr\"{o}m-type discretization ---
the density is approximated 
by its values at the discretization nodes,
which we denote by
$\sigma_{qp} \approx \sigma(\xx_q(t_p))$.
%
The basis of a GGQ rule is a set of 
support nodes and weights for the
contribution to the integral from the
``self'' panel (panel $j$) and the adjacent
panels (with index in $A(j)$).
%

@@

@@

@@

Below, we let $D^N_k$, $S^N_k$ and $W^N$ denote
the discretized versions of the $\cDk$,
$\cSk$, and $\cW$ operators, respectively.

@@

>>>>

\subsection{Numerical methods}

\subsubsection{Discretizing the BIE}

@@

Let the boundary be divided into $N_p$
panels.
%
We parameterize panel $j$ as
$\bx_j(t)$, with $t$ ranging over the
interval $[-1,1]$.
%
Each component of $\bx_j$ is taken to be
a polynomial interpolant over the
standard 16th-order Legendre nodes on
$[-1,1]$, denoted by $t_n$, so that
the total number of discretization
points is $N=16N_p$.
%
See \cref{fig:panels} for an example
discretization.
%
Another important quantity below is the
arc-length density of a panel, which
we denote by $s_j(t) := |\bx'_j(t)|$.
%
Finally, we denote the set of panels
which are adjacent to panel $j$
by $A(j)$. On a closed curve,
$A(j)$ contains two integers.

The integral kernels of the single and
double layer potentials have weak
singularities of the form $r^p\log r$
which require special quadrature rules to
achieve high-order accuracy.
%
In the examples below, we use generalized
Gaussian quadrature (GGQ)~\cite{bremer2010}.
%
To demonstrate the idea, we consider
evaluating the convolution of a kernel
$K$ with a density $\sigma$
at the boundary node $\xx_j(t_l)$.
%
GGQ is a Nystr\"{o}m-type discretization ---
the density is approximated 
by its values at the discretization nodes,
which we denote by
$\sigma_{qp} := \sigma(\xx_q(t_p))$.
%
The basis of a GGQ rule is a set of 
support nodes and weights for the
contribution to the integral from the
``self'' panel (panel $j$) and the adjacent
panels (with index in $A(j)$).
%

@@

\begin{remark}
  \label{rmk:levelrestrict}
  We ensure that ``far'' interactions
  are handled to high precision by requiring that
  no two adjacent panels differ in length
  by more than a factor of 2. On a domain which does
  not nearly self-intersect this
  guarantees that no ``far'' interactions occur
  which are much closer than 1/2 of a panel away
  (assuming panels are relatively flat).
  Because the location of the singularity is
  bounded away from the panel and the smooth
  rule is of high order, we obtain a quadrature
  rule with sufficient precision.
\end{remark}

\subsubsection{Fast determinant method}

@@

The compressed representation also allows
for fast applications of the system matrix,
its transpose, the inverse of the system
matrix, and the inverse transpose to
vectors.
%
In particular, this allows us to estimate the
smallest singular values by performing
randomized subspace iteration, see
\cite[Algorithm 4.4]{halko2011finding},
on the inverse operator.
%
Below, we use the smallest singular value
as a measure of the quality of the
eigenvalues found by approximating the
roots of the determinant.
%
We also evaluate the second smallest singular
value if the root finding precedure suggests a
possible double root.

\subsubsection{Interpolation and root-finding}

@@

The \texttt{roots} utility returns the roots
of the polynomial in the complex plane, with
some minimal internal processing to remove
spurious roots.
%
Because our numerical determinant evaluation
is somewhat noisy and we fit the function up to
precision $\epscheb$, we perform some
further post-processing to eliminate remaining
spurious roots.
%
We ignore any of the returned roots $z$ with $\imag(z)
> \sqrt{\epscheb}$, as these are too far from real-valued
to be non-spurious.
%
For the remaining roots, we inspect any
pairs of roots which are closer than
$\epsclose = 10^{-5}$.
%
For these roots, we compare the resulting
eigenfunctions evaluated at a large number of points
within the domain.
%
If the eigenfunctions are the same to
reasonable precision, we declare that these
roots correspond to a single root and
retain the root for which the operator
has the smallest singular value.
%
For these near double-roots, we also check
that there is no two dimensional null-space
corresponding to the root by estimating the
second smallest singular value.

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{fig/barbell_gallery}
  \caption{Vorticity plots of the first 119 eigenfunctions
    of the barbell-shaped domain.}
  \label{fig:barbell_gallery}
\end{figure}

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.4\textwidth}
    \centering
    \includegraphics[width=\textwidth]{fig/ex_barbell_001_sings_plot}
    \caption{Smallest singular value corresponding to the
    first 119 computed eigenfunctions of the barbell-shaped
    domain.}
    \label{subfig:barbell_sings}
  \end{subfigure}
  ~
  \begin{subfigure}[t]{0.4\textwidth}
    \centering
    \includegraphics[width=\textwidth]{fig/ex_barbell_001_coeffs_plot}
    \caption{Normalized Chebyshev coefficients of $f^N(k)$ on
      3 different intervals in $k$.}
    \label{subfig:barbell_coeffs}
  \end{subfigure}
  \caption{Diagnostics for the barbell eigenvalues.}
  \label{fig:barbell_diagnostics}
\end{figure}

We consider the barbell-shaped domain in \cref{fig:barbell}.
This domain is the union of a square of side-length 6,
a square of side-length 3, and a ``bridge'' connecting
them of height 1 and width 5/2.
%
For simplicity, we round the corners of the domain
to obtain a smooth object.
%
Applying the approach described in~\cite{epstein2016smoothed},
the corners of the domain are rounded using the Gaussian kernel
\begin{equation}
  \nonumber
\phi(x) = \frac{1}{\sqrt{2\pi h}} e^{-x^2/(2 h^2)} \, ,
\end{equation}
with $h\approx 0.06$. This leaves the domain unperturbed
outside of a radius of $0.1$ around each corner.
%
The eigenfunctions of such a domain display the well-known
localization property~\cite{trefethen2006computed}:
many of the eigenfunctions are approximately supported
within one of the squares.
%
We compute these eigenfunctions corresponding to
eigenvalues $k^2$ with $k$ in the range
$0.5 \leq k \leq 6.5$.

%
The panels are chosen adaptively so that the coordinates,
the first and second derivatives of the coordinates,
and the arc-length function are resolved to a precision
of $10^{-12}$ on each panel.
%
This results in $N_p = 412$ after enforcing the
level-restriction property described in
\cref{rmk:levelrestrict}
and enforcing that no panel is larger than
one wavelength for the largest $k$
(here $\lambda=2\pi/6.5$).

We compute the eigenvalues as the values of
$k$ for which $\cI-2\cDk-2\cW$ is non-invertible.
%
Let $f^N(k) = \det (\cI^N-2\cDk^N-2\cW^N)$.
To find the roots of $f^N(k)$, we fit a \texttt{chebfun}
representation of $f^N(k)$ on each of the intervals
$[j/2,(j+1)/2]$ for $j = 1,\ldots,12$.
%
We plot the absolute value of the Chebyshev coefficients
of $f^N(k)$ on the intervals $[0.5,1.0]$, $[3.0,3.5]$,
and $[6.0,6.5]$ in \cref{subfig:barbell_coeffs}.
%
As expected, the coefficients decay exponentially
to zero, with more terms required at higher
frequencies.
%

We compute the roots of these Chebyshev interpolants
and apply the post-processing described above.
%
There were 135 total roots: 2 were removed because
$\Im(z) > \sqrt{\epscheb}$ and 13 pairs were found
with values within $\epsclose = 10^{-5}$ of each other.
%
For these 13 pairs, none corresponded to a double
root.
%
This leaves 119 roots in the range $0\leq k \leq 6.5$.
%
We plot the smallest singular value of
$\cI^N-\cDk^N-2\cW^N$ for each of these roots in
\cref{subfig:barbell_sings}
and plot the vorticity of the eigenfunctions
in \cref{fig:barbell_gallery}.
%
The singular values suggest that the quality of the
eigenvalues is good.
From the plots, we see that localization occurs
until about the 100th eigenvalue.

%
